---
title: 'Modelling'
output:
    html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r, message=FALSE, echo=FALSE, warning=FALSE}
library(data.table)
library(tidyverse)
library(knitr)
library(ggplot2)
library(caret)
library(mgcv)
library(car)
library(leaps)
library(MASS)
library(rpart)
library(rpart.plot)
library(randomForest)
library(gbm)
library(xgboost)
library(doParallel)
library(plotly)
```

```{css, echo = FALSE}
caption {
      color: black;
    }
```

```{r}
restaurants <- read.csv("data/cleaned_restaurants.csv") # from `data-wrangling.Rmd`
```

```{r}
restaurants <- restaurants %>% 
  mutate(
    price_level = if_else(is.na(price_level), "Unknown", price_level)) %>% 
  mutate (
    price_level = factor(price_level, 
                         levels = c("Low", "Medium", "High", "Very High", "Unknown")),
    size_level = factor(size_level, 
                         levels = c("Single Location", "Small Chain", 
                                    "Medium Chain", "Large Chain")),
    type = factor(type),
    cuisine = factor(cuisine)
  )
```

```{r}
set.seed(9900)
train <- sample(1:nrow(restaurants), round(0.7 * nrow(restaurants)))

rest_train <- restaurants[train, ]
rest_test <- restaurants[-train, ]
```

# Overall Regression Results
**Table 1** suggests that the gradient boosting model had the best overall performance, with a somewhat low train RMSE and the lowest test RMSE. The random forest model had the lowest train RMSE, but a higher test RMSE, suggesting possible overfitting.
```{r}
data <- data.frame(
  Method = c("Linear Regression", "Auto. Lin. Reg.", "GLMM", "Regression Tree", 
             "Random Forest", "Boosting", "XGBoost"),
  Train_RMSE = c(0.634, 0.634, 0.62, 0.64, 0.323, 0.597, 0.545),
  Test_RMSE = c(0.634, 0.634, 0.63, 0.633, 0.635, 0.626, 0.63)
)

kable(data, format = "pipe", caption = "**Table 1**. RMSE results for each model", col.names = c("Method", "Train RMSE", "Test RMSE"))
```

# Variable Importances
**Figure 1** shows the rankings of each variable in the random forest, boosting and XGBoost models. There is a lot of variations between the rankings, but overall, size level appears to be the most important.
```{r}
importances <- data.frame(
  model = rep(c("Random Forest", "Boosting", "XGBoost"), each = 8),
  variable = rep(c("latitude", "longitude", "distance", 
                   "review_count", "size_level", "cuisine", 
                   "type", "price_level"), times = 3),
  rank = c(1, 2, 3, 4, 5, 6, 7, 8,
           6, 7, 8, 3, 1, 4, 5, 2,
           3, 5, 4, 6, 1, 7, 8, 2)
)

importances$model <- factor(importances$model, levels = c("Random Forest", "Boosting", "XGBoost"))
```

```{r diff, fig.cap="**Figure 1.** Differences in Variable Importance Rankings Over Models"}
library(ggbump)

imp <- ggplot(importances, aes(x = model, y = rank, group = variable, 
                               color = variable)) +
  geom_bump(size = 1.5, smooth = 10) +
  geom_point(size = 4) +
  scale_y_reverse(breaks = 1:25) +  # 1st at top
  theme_minimal() +
  labs(x = "Model",
       y = "Rank") +
  theme(legend.position = "bottom")

ggplotly(imp)
```